The Eclipse Integrated Computational Environment

### 2018-07-12 src/review-response-20180712.tex

Dear Dr. Sobie,

Thank you very much for providing the reviewer's responses. Please find below a list of the issues that we have addressed and comments on the reviewer's responses. We found these reviews to be highly constructive and made every effort to address the reviews as thoroughly as possible. We believe that the result is a significantly stronger narrative.

@@

We have resubmitted the manuscript and hope that you will reconsider it for publication in Software X.

Best,
Jay Jay Billings
ORNL

-----
Reviewer #1
>> 1) A figure of some sort on complete architecture is extremely desired.

>> 2) Some of the references need to be looked up carefully. For instance reference 23 and 24 seem incomplete and less useful for a reader. See below:
>> 
>> [24] Jay Jay Billings. Eclipse ICE, October 2016.

>> 3) Renaming Comparison to other models seems incomplete. Perhaps it could be renamed to something better such as "related work."

@@

We have expanded this section and addressed unique features and the comparison of ICE to other workflow engines with activities-based workflow models.

>> 4) It is not very clear exactly how ICE integrate code from other languages. A reference to EASE was given but no other detail was provided.

We have added text to clarify that this is done by simple scripts in Javascript and Python through a built in shell, through workflow items, and in extreme cases through the Java Native Interface.

-----
Reviewer #2:

>> 1) Related work should be carefully revised, and include popular workflow system such as Swift, Pegasus, Makeflow, Taverna, Askalon, Nextflow, etc. (only Kepler is shown as an example). The manuscript claims that ICE follows the five categories described in the motivation section, however it is not the only workflow system that follows this classification.

@@

We do not believe that a detailed, n-by-n comparison of workflow systems is valuable in the context of a software publication on ICE, but we agree that it is important to illustrate the differences a little better. To that end, the text now discusses Chiron, Taverna, Aiida, and Fireworks. We have also added a reference to other work by the authors that describes the differences between modeling and simulation workflows and grid workflows in more detail.

>> 1.b) The same issue appears again in Section 2.0.2 (section number should be revised too), where the presented model is claimed to significantly differ from similar efforts in workflow science. This is not accurate, there are several workflow systems that follows the activities model, such as Moteur, Taverna, and Chiron. It is crucial to distinguish ICE's contributions from these systems.

@@

Taverna and Chiron are now discussed in the related work section. The exact differences between their Activities and ICE's are presented.

>> 2) Last, interactive analysis of workflow executions is claimed as a novelty of this work; however workflow steering has been investigated by over a decade and support in systems such as Chiron and SciCumulus. How ICE differs from them?

Our paper only points out that ICE can perform this work and does not, as far as we can tell, claim that ICE uniquely provides this capability.  We did not address anything related to this point, but would be happy to do so with more direction if necessary.

>> 3) Section 3, software description needs to be substantially improved. It misses an overview architecture (a figure would be extremely helpful) that shows how ICE components interact/leverage Eclipse components. This is one of the main contributions of this paper and should be better articulated.

@@

We have addressed this point by adding new diagrams and more text describing the architecture of ICE. We have also added more information on the relationship to Eclipse Equinox, RCP, and PTP, etc. and cleaned up the sections that describe Eclipse across the document.

>> 4) How the Items model differs from Taverna's model with Web Services?

@@

Taverna's web services are now discussed specifically in the related work section, but since they are not similar to ICE's Items the relationship is not addressed. Shim services are now discussed briefly as well.

>> 5) There is a mention that ICE API will be evolved to make the API language closer to other systems. Which systems and how?

@@

This is a generic statement about the names of classes and methods in other systems compared to the language used to describe these in ICE. For example, Item's in ICE are typically called tasks in Kepler-based systems such as Triquetrum.

The footnote has been changed to reflect this by adding "such as Triquetrum" after "other system" to provide an example.

>> 6) What defines a small output (so it will be automatically transferred)?

@@

50MB. This has been updated in the text.

>> 7) Why is it more sophisticated? What can be achieved here? It is important to enumerate the advantages.

We have updated the text in this section to more clearly and fairly reflect our thinking on language-based workflows in ICE.

>> 8) What is a 'stubs' of plugins? How will it then interact with the framework? Which capabilities are used to enable them with ICE?

We have updated the text to define stubs, strengthen the related statement on installation, and to describe the role of the OSGi in this process.

>> 9) Which operations are available via the APIs? It is mentioned that it has a rich API, but no detailed information is provided.

@@

Detailed information is provided in the API documentation, which is cited in the final section. However, we strengthened this sentence by adding text that all workflow management functions are exposed by the API.

>> 10) IMHO, one of the main contributions and novelty of the paper is the combination of the Eclipse IDE with the workflow development/execution capabilities. However, it is not discussed in detail how a user may benefit from this integration.

This point has been addressed by creating a section specifically on the relationship to Eclipse by merging content from the old framework extension section and new information on the benefits of self-hosting.

>> 11) The examples section would significantly benefit of more information about how the workflows are executed.

We added a paragraph describing how workflows are executed in ICE at the end of the executing jobs section. The main purpose of this paragraph was to illustrate the differences between executing jobs and executing workflows because they are significantly different.

>> 12) Also, there are mentions that ICE can run locally and in HPC platforms, but no information is provided on how the system actually executes the workflows. There is a quick mention to SSH connections, but nowadays there are other methods to remotely submit jobs for execution, such as Globus GRAM, Bosco, etc. The manuscript needs a proper workflow execution section which describes the different methods how the workflows can be executed.

@@

The executing jobs section completely covers how workflows are executed in ICE. We have modified it to state that Globus GRAM, etc. are not supported. We have also provided some text to clarify that executing jobs is not the same as executing a workflow.

>> 13) In summary, I see much potential in the proposed workflow tool, however a thoroughly characterization and description of the architectural components and their capabilities is key to demonstrate the novelty and impact of the discussed system.

We are very grateful for this thorough review of the paper. We believe that we have addressed this review just as thoroughly with the significant updates provided in this revised manuscript, especially the new content on the architecture and the relationship to other Eclipse technologies.

### 2018-07-13 src/content.tex
Addressed editor's comments. Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

Computational scientists perform a variety of tasks in modeling and
simulation that can be abstracted into a lightweight theoretical
framework based broadly on five high-level activities: (1)
creating input, (2) executing jobs, (3) analyzing results, (4) managing
data, and (5) modifying code. Those same computational scientists would
most likely find these activities difficult for all codes with which
they lack experience, whereas with their own codes---or those with which
they are most familiar---these tasks may be so simple that they are
taken for granted. Any particular combination of these activities across
one or more scientific software package or code results in a unique
workflow. Such a workflow is normally, but not always, requested
by a human user and orchestrated by a workflow management system.

@@

Although this is similar to executing a job, it is distinctly
different because the activity changes focus to satisfy the needs of a
human operator. Simple data reduction, where the exact reduction is
known, certainly qualifies as executing a job; however, analysis of
modeling and simulation results is far from simple data reduction and is
generally far more interactive for computational scientists.

@@

\textbf{Modifying code} is not typically considered a part of a
scientific computing workflow. However, modeling and simulation use
cases often require users to explicitly modify code before execution or to issue special re-build instructions. This is true, for example, with the computational fluid dynamics code Nek5000, which requires modification of code before compilation and special build instructions using the \textit{makenek} shell script
\cite{the_nek5000_team_nek5000_2014}.
Many computational scientists consider ``their workflow'' to be re-running
software after modifications for purely exploratory purposes. This might
be required if the model that the author is modifying cannot be
configured directly as part of the input but can be easily accomplished by manipulating the source code.

@@

ICE was specifically created to address hands-on workflows for computational scientists, as in Section \ref{illustrative-examples}. Users download and
execute ICE locally, and ICE in turn orchestrates local or remote workflows as
required. It provides a comprehensive workbench for modeling and
simulation that includes tools for workflows, visualization, data
management, and software development.

@@

The most important function of ICE is to serve as an easily extended
workflow management system for computational scientists in support of
the activities described previously. In practice, it is most often used as a
combination of workflow management system and development environment
since it contains a significant amount of Eclipse's software development
tooling in addition to the workflow tools. The project has many users,
but it is most heavily used by the development team to support workflow
science and software development for energy science projects. The
development team regularly uses the platform to quickly deploy new
domain-specific workbenches in a matter of hours for small collections
of workflows that are easy to encode.

@@

Modeling and simulation workflows for computational scientists differ
greatly from those of experimentalists or those who primarily interact with
grid-based workflow management systems. The Eclipse ICE effort described here 
has been used to address interdisciplinary problems in 
modeling and simulation for energy science. Of particular interest are the
differences in architecture between a workflow management system focused
on modeling and simulation compared with those focused on grids. ICE's broad
applicability across many topics in energy science suggests that
there are opportunities for these systems in general. Finally, one 
interesting avenue of future exploration is coupling or integrating ICE 
with other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus,
which would make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.

>>>>

ICE's workflow model is based on making it easier for scientists to create input, launch jobs, analyze results, manage data, and modify code. Many scientists would
most likely find these activities difficult for all codes with which
they lack experience, whereas with their own codes---or those with which
they are most familiar---these tasks may be so simple that they are
taken for granted. Any particular combination of these activities across
one or more scientific software package or code results in a unique
workflow. Such a workflow is normally, but not always, requested
by a human user and orchestrated by a workflow management system.

@@

Although this is similar to executing a job, it is distinctly
different because the activity changes focus to satisfy the needs of a
human operator. Simple data reduction, where the exact reduction is
known, certainly qualifies as executing a job; however, analysis of
modeling and simulation results is far from simple data reduction and is
generally far more interactive for scientists.

@@

\textbf{Modifying code} is not typically considered a part of a
scientific computing workflow. However, modeling and simulation use
cases often require users to explicitly modify code before execution or to issue special re-build instructions. This is true, for example, with the computational fluid dynamics code Nek5000, which requires modification of code before compilation and special build instructions using the \textit{makenek} shell script
\cite{the_nek5000_team_nek5000_2014}.
Many scientists consider ``their workflow'' to be re-running
software after modifications for purely exploratory purposes. This might
be required if the model that the author is modifying cannot be
configured directly as part of the input but can be easily accomplished by manipulating the source code.

@@

ICE was specifically created to address hands-on workflows for scientists, as in Section \ref{illustrative-examples}. Users download and
execute ICE locally, and ICE in turn orchestrates local or remote workflows as
required. It provides a comprehensive workbench for modeling and
simulation that includes tools for workflows, visualization, data
management, and software development.

@@

The most important function of ICE is to serve as an easily extended
workflow management system for scientists in support of
the activities described previously. In practice, it is most often used as a
combination of workflow management system and development environment
since it contains a significant amount of Eclipse's software development
tooling in addition to the workflow tools. The project has many users,
but it is most heavily used by the development team to support workflow
science and software development for energy science projects. The
development team regularly uses the platform to quickly deploy new
domain-specific workbenches in a matter of hours for small collections
of workflows that are easy to encode.

@@

Modeling and simulation workflows for scientists differ
greatly from those of experimentalists or those who primarily interact with
grid-based workflow management systems. The Eclipse ICE effort described here 
has been used to address interdisciplinary problems in 
modeling and simulation for energy science. Of particular interest are the
differences in architecture between a workflow management system focused
on modeling and simulation compared with those focused on grids. ICE's broad
applicability across many topics in energy science suggests that
there are opportunities for these systems in general. Finally, one 
interesting avenue of future exploration is coupling or integrating ICE 
with other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus,
which would make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.

### 2018-07-13 src/main.tex
Fixed a typo in the abstract. Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>
Addressed editor's comments. Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

Problems in modeling and simulation require significantly different
workflow management technologies from standard grid-based workflow
management systems. Computational scientists typically interact with
simulation software in a feedback-driven way where solutions and
workflows are developed iteratively and simultaneously. This work
describes common modeling and simulation activities in workflows and how combinations of these
activities form unique workflows. It presents the Eclipse Integrated
Computational Environment as a workflow management system and
development environment for the modeling and simulation community.
Examples of the Environment's applicability to problems in energy
science, general multiphysics simulations, quantum computing, and other
areas are presented along with its impact on the community at large.

@@

\section{Notice of Copyright}\label{notice-of-copyright}

>>>>

Problems in modeling and simulation require significantly different
workflow management technologies from standard grid-based workflow
management systems. Computational scientists typically interact with
simulation software in a feedback-driven way where solutions and
workflows are developed iteratively and simultaneously. This work
describes common modeling and simulation activities and how combinations of these
activities form unique workflows. We present the Eclipse Integrated
Computational Environment as a workflow management system and
development environment for the modeling and simulation community.
Examples of the Environment's applicability to problems in energy
science, general multiphysics simulations, quantum computing, and other
areas are presented along with its impact on the community at large.

@@

\section*{Notice of Copyright}
